Manuscript_Github_Pchabaudi_calculateTPM

### 2022-07-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_calculateTPM.Rmd
Change more names to chab_tpm base and save pseudogene image

#Do they all sum to 1e6? Note that all.equal allows for a tiny bit of difference which is okay.
all.equal(apply(twentyfour_ct_comb[-1], 2, sum), 
          rep(1e6, times = dim(twentyfour_ct_comb[-1])[2]), 
          check.names = FALSE)
```

@@

#Setting twentyfour_tpm_ct core dataset - remove pseudogene and re-scaled

twentyfour_tpm_nopseudo <- chab_tpm.df %>% 
  filter(!Geneid %in% 'PCHAS_0420800')

twentyfour_tpm_rescale_nopseudo <- lapply(colnames(twentyfour_tpm_nopseudo), 
                                          function(column_name){
                                            column <- getElement(twentyfour_tpm_nopseudo, 
                                                                 paste(column_name))
                                            if(is.numeric(column)){
                                              #is.numeric to avoid trying to scale the Geneid column
                                              tpm_scale_factor <- 1e6 / sum(column)
                                              tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                            }else{
                                              tpm_scaled <- column
                                            }
                                            return(tpm_scaled)
                                          }) %>% 
  bind_cols

colnames(twentyfour_tpm_rescale_nopseudo) <- colnames(twentyfour_tpm_nopseudo)
# apply(twentyfour_tpm_rescale_nopseudo[,-1], 2, sum)

#Setting the re-scaled version as the default from now on
#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
chab_tpm.df <- twentyfour_tpm_rescale_nopseudo
```

```{r clean_up_and_average}
# Change the names to remove superfluous 'rMT'

chab_tpm.df <- rename_with(chab_tpm.df, 
                                  ~ str_extract(.x,pattern = '(?<=rMT_).+'),
                                  matches('rMT'))

twentyfour_rMT_avg <- melt(chab_tpm.df, 
                           variable.name = 'sample',
                           value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
  group_by(Geneid, stage) %>% 
  summarise(mean_tpm = mean(tpm)) %>% 
  dcast(Geneid ~ stage)
```

>>>>

#Do they all sum to 1e6? Note that all.equal allows for a tiny bit of difference which is okay.
all.equal(apply(chab_tpm.df[-1], 2, sum), 
          rep(1e6, times = dim(chab_tpm.df[-1])[2]), 
          check.names = FALSE)
```

```{r super-pseudogene}

chab_tpm_melt.df <- reshape2::melt(chab_tpm.df,
                                   variable.name = 'sample',
                                   value.name = 'tpm') %>% 
  mutate(time = str_extract(sample, 
                            pattern = '([[:digit:]]{2}h_.+|sch.culture)(?=_[[:digit:]])'),
         transmission = str_extract(sample,
                                    pattern = 'rMT|SBP'))

(plot <- ggplot(data = filter(chab_tpm_melt.df, 
                     Geneid %in% 'PCHAS_0420800'), 
       aes(x = time, 
           y = log1p(tpm),
           fill = transmission)) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
               geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
  geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
  scale_fill_manual(values = cbPalette) +
  theme_classic() +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)))

ggsave(filename = 'plots/pseudogene_tpm_barchart.png',
       plot)

@@

#Setting chab_tpm_ct core dataset - remove pseudogene and re-scaled

chab_tpm_nopseudo <- chab_tpm.df %>% 
  filter(!Geneid %in% 'PCHAS_0420800')

chab_tpm_rescale_nopseudo <- lapply(colnames(chab_tpm_nopseudo), 
                                          function(column_name){
                                            column <- getElement(chab_tpm_nopseudo, 
                                                                 paste(column_name))
                                            if(is.numeric(column)){
                                              #is.numeric to avoid trying to scale the Geneid column
                                              tpm_scale_factor <- 1e6 / sum(column)
                                              tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                            }else{
                                              tpm_scaled <- column
                                            }
                                            return(tpm_scaled)
                                          }) %>% 
  bind_cols

colnames(chab_tpm_rescale_nopseudo) <- colnames(chab_tpm_nopseudo)
# apply(chab_tpm_rescale_nopseudo[,-1], 2, sum)

#Setting the re-scaled version as the default from now on
#Setting chab_tpm_ct core dataset - removed pseudogene and re-scaled
# chab_tpm.df <- chab_tpm_rescale_nopseudo
```

```{r clean_up_and_average}
chab_tpm_avg.df <- reshape2::melt(chab_tpm.df, 
                           variable.name = 'sample',
                           value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
  group_by(Geneid, stage) %>% 
  summarise(mean_tpm = mean(tpm)) %>% 
  reshape2::dcast(Geneid ~ stage)
```
